---
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# SI 618 Homework 7

# Part 1 (40 points)

## Question 1: Load country data (5 points)

First the provided TSV data file is loaded into R using the **read.table()** function. Display the first 15 rows
of the data frame:

```{r echo=FALSE}
data <- read.table(file="countrydata_withregion.tsv", sep="\t",quote="", header=TRUE)
data[1:15,1:4]
```

## Question 2: Scatter plot of log transformed data (5 points)

Logarithms (base 2) of the area and the population of each country are computed and used to produce the
following scatter plot using the **qplot()** function. Use **{r echo=FALSE, fig.width=7}** for all of the plots.

```{r echo=FALSE, fig.width=7}
library(ggplot2)
qplot(log2(area), log2(population), data=data, xlab="Log base 2 of the area", ylab="Log base 2 of the population")
```

## Question 3: Data aggregation by region (15 points)

The areas and populations of all countries in a region are summed up using the **aggregate()** function,
respectively. Then the following two pie charts are created using the **pie()** function.

```{r echo=FALSE, fig.width=7}
agg = aggregate(list(data$area, data$population), by=list(data$region), FUN=sum, na.rm=TRUE)
pie(agg[,2],labels=agg[,1],main="Area of Regions")
```

```{r echo=FALSE, fig.width=7}
pie(agg[,3],labels=agg[,1],main="Population of Regions")
```

## Question 4: Visualization of Population per sq km of Regions (15 points)

A new data frame is created to contain the population per sq km of each region using the **data.frame()**
function. The data frame is then sorted by population per sq km in decreasing order with the help of the
**reorder()** function. Finally, the following bar plot is created using the **qplot()** function with **geom=“bar”**.
In order to rotate the x-axis labels, add **+ theme(axis.text.x = element_text(angle = 60, hjust =
1))** at the end of the **qplot()** function call.

```{r echo=FALSE, fig.width=7}
population_per_sq = data.frame(region = agg[,1],
                               value = agg[,3]/agg[,2])
population_per_sq$region <- reorder(population_per_sq$region, X=-population_per_sq$value, order=is.ordered(population_per_sq))
qplot(region,value,data=population_per_sq,geom = "col",ylab="Population per sq km of Regions", xlab="Region",)+ theme(axis.text.x = element_text(angle = 60, hjust = 1))
```


# Part 2 (60 points)
## Question 5: Load yelp data & generate summary (10 points)

Load the TSV data file: businessdata.tsv into a R data frame using the **read.table()** function. The city, state
and main_category columns should be converted to factors. Then listwise deletion (http://en.wikipedia.org/wiki/Listwise_deletion) is applied to remove records with missing data (use the **na.omit()** function). Then
the data.frame is converted to a data.table. Here is the summary of the data table:

```{r echo=FALSE}
data <- read.table(file="businessdata.tsv", sep="\t", header=TRUE,fill=TRUE,quote="", comment.char="", colClasses=c("character","factor","factor","numeric","numeric","factor"))
df <- data
data <- na.omit(data)
summary(data)
```

## Question 6: Histogram of Star Rating (10 points)

The Histogram of star ratings is plotted with the **qplot()** or **ggplot()** function. The actual counts plot is
shown. (Use **binwidth=1**)

```{r echo=FALSE, fig.width=7}
qplot(data=data, mapping=aes(x=stars, fill=state),geom="histogram",binwidth=1,xlab="Stars",main="Histograms of Star Ratings")+ facet_grid(. ~ state)
```

## Question 7: Histograms of Review Counts (10 points)

Histograms of review counts are plotted with the **qplot()** or **ggplot()** function. (Use **binwidth=10**)

```{r echo=FALSE, fig.width=7}
qplot(data=data, mapping=aes(x=review_count),geom="histogram",binwidth=10,xlab="Review Counts",main="Histograms of Review Counts", alpha=0.8, show.legend = F)
```

We can see that the distribution of review counts has a long tail. To zoom in on the bars to the left of the
200 mark, we use the **data.table** syntax or the **subset()** function to select just the data with review count
<= 200. And then plot the histogram again with **binwidth=1**.

```{r echo=FALSE, fig.width=7}
qplot(data=data[data$review_count<=200,], mapping=aes(x=review_count),geom="histogram",binwidth=1,xlab="Review Counts",main="Histograms of Review Counts (Filtered)", alpha=0.8, show.legend = F)
```

## Question 8: Boxplot of Star Ratings by States (10 points)

```{r echo=FALSE, fig.width=7}
qplot(data=data, mapping=aes(x=state,y=stars,fill=state),geom="boxplot",ylab="Stars",main="Star Ratings by States")
```

## Question 9: Bar Chart of Number of Businesses by State (10 points)

The states should be orderd by descreasing height of bars. Use the **reorder()** function.

```{r echo=FALSE, fig.width=7}
data$state <- reorder(data$state,X=data$state, FUN=length)
data$state <- factor(data$state, levels=rev(levels(data$state)))
qplot(data=data,mapping=aes(x=state),geom="bar",ylab="State",alpha=0.8,show.legend=F)
```
